Stability and Chaos in Multi-Agent Reinforcement Learning

### 2020-10-08 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
minor mods

We note that the numerical experiments confirm the predictions suggested by the analytic results shown in Figure \ref{sec::Theory}. We notice that convergence occurs almost only for low values of $\alpha$. This observation remains as we increase the value of $N$. As noted, the analytic result over-estimates the region of instability due to the assumptions of $N \rightarrow \infty$ and $\frac{1}{p-1} << 2$ made during the derivation. Remaining discrepancies are partially due to the aforementioned approximation in the calculation of $q$, $\chi$ and the fact that the analytic result considers a continuous time approximation of the expected behaviour of Q-Learning. We expect that testing over a greater number of payoff realisations and initial conditions will yield a more representative assessment of the average behaviour of Q-Learning. However, running these experiments is a computationally expensive procedure: for $p$ players and $N$ actions we require operations on $p$ matrices with $N^{p}$ elements. As such, due to a reduced availability of computational facilities, a large scale averaging was not possible.

However, due to the abundance of non-convergent behaviour, shown by the dark regions in Figure \ref{fig:NumericalExperiments} other than at low values of $\alpha, \tau$, we can conclude that the implications 1, 2 are verified. Indeed implications 3 and 4 are also verified as the proportion of games converged decrease monotonically with $\alpha, \tau$, but there does not appear to be any discernible dependence on $\Gamma$. Finally, we can conclude that implication 5 is verified by comparing Figures \ref{fig:pvariation} and \ref{fig:pVarianceExpt}. It is clear that the instability of the system increases monotonically with $p$. A point to note is that Figure \ref{fig:pVarianceExpt} shows a logarithmic dependence of the instability on $p$, whilst Figure \ref{fig:pvariation} suggests that the dependence should be exponential. Due to the limited range of $p$ that was used to generate Figure \ref{fig:pVarianceExpt}, it cannot yet be concluded that the dependence on $p$ is logarithmic. However, the important result is that the unstable behaviour increases, regardless of parameter choice as $p$ increases, which is corroborated by both theory and experiment. This phenomenon was referred to by Sanders et. al as `the prevalence of chaotic dynamics in games with many players' \cite{Sanders2018}.

@@

To summarise, we have shown that the analytic result (\ref{eqn::Final}) provides a strong assessment for the effect that $\alpha$, $\tau$, $\Gamma$ and $p$ have on the likelihood of convergence of Q-Learning. We show that, at least for negatively correlated payoffs, the strength of correlation plays no role in determining the stability of the system. Furthermore, though (\ref{eqn::Final}) overestimates the region of instability, it accurately conveys that convergence of Q-Learning is unlikely for games with many players and actions. In fact, our experimental results also verify that stability may only be guaranteed for the limiting case of two-player, two-action games.

Finally, we show that the behaviour of the stability line differs from that of EWA and suggest that this shows that a stability analysis is an important mode of analysis for reinforcement learning equations to ensure that parameters are being chosen appropriately to guarantee the safe convergence of learning.

>>>>

We note that the numerical experiments confirm the predictions
suggested by the analytic results shown in Figure \ref{sec::Theory}
\fb{but Figure \ref{sec::Theory} concerns the numerical
  experiments}. We notice that convergence occurs almost only for low
values of $\alpha$. This observation remains as we increase the value
of $N$. As noted, the analytic result over-estimates the region of
instability due to the assumptions of $N \rightarrow \infty$ and
$\frac{1}{p-1} << 2$ made during the derivation. Remaining
discrepancies are partially due to the aforementioned approximation in
the calculation of $q$, $\chi$, and the fact that the analytic result
considers a continuous-time approximation of the expected behaviour of
Q-Learning. We expect that testing over a greater number of payoff
realisations and initial conditions will yield a more representative
assessment of the average behaviour of Q-Learning. However, running
these experiments is a computationally expensive procedure: for $p$
players and $N$ actions it requires operations on $p$ matrices with
$N^{p}$ elements. As such, due to a reduced availability of
computational facilities, a large scale averaging was not possible.

However, due to the abundance of non-convergent behaviour, shown by
the dark regions in Figure \ref{fig:NumericalExperiments} other than
at low values of $\alpha, \tau$, we can conclude that the implications
(1), (2) in Sec.~\ref{} are verified. Indeed implications (3) and (4) are also verified as
the proportion of converging games
%converged
decreases monotonically with $\alpha,
\tau$, but there does not appear to be any discernible dependence on
$\Gamma$. Finally, we can conclude that implication (5) is verified by
comparing Figures \ref{fig:pvariation} and \ref{fig:pVarianceExpt}: it
is clear that the instability of the system increases monotonically
with $p$. A point to note is that Figure \ref{fig:pVarianceExpt} shows
a logarithmic dependence of the instability on $p$, whilst Figure
\ref{fig:pvariation} suggests that the dependence should be
exponential. Due to the limited range of $p$ that was used to generate
Figure \ref{fig:pVarianceExpt}, it cannot yet be concluded that the
dependence on $p$ is logarithmic. However, the important result is
that the unstable behaviour increases, regardless of parameter choice
as $p$ increases, which is corroborated by both theory and
experiment. This phenomenon was referred to by Sanders et. al as `the
prevalence of chaotic dynamics in games with many players'
\cite{Sanders2018}.

@@

To summarise, we have shown that the analytic result
(\ref{eqn::Final}) provides a strong assessment for the effect that
$\alpha$, $\tau$, $\Gamma$ and $p$ have on the likelihood of
convergence of Q-Learning. We showed that, at least for negatively
correlated payoffs, the strength of correlation plays no role in
determining the stability of the system. Furthermore, though
(\ref{eqn::Final}) overestimates the region of instability, it
accurately conveys that convergence of Q-Learning is unlikely for
games with many players and actions. In fact, our experimental results
also verify that stability may only be guaranteed for the limiting
case of two-player, two-action games \fb{actually we consider 3
  players as well}.
%
Finally, we showed that the behaviour of the stability line differs
from that of EWA and suggest that this shows that a stability analysis
is an important mode of analysis for reinforcement learning algorithms
to ensure that parameters are being chosen appropriately to guarantee
the safe convergence of learning.
